Davos: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-09-18 paper/main.tex
edits to section 2.2.4 up through writeable options

@@

Davos's behavior may be customized by modifying a set of attributes attached to
the \texttt{davos} module object that is added to the workspace when Davos is
imported. These attributes may be modified, displayed, or checked
programmatically at runtime (see Sec.~\ref{sec:illustrative-example} for an
illustrative example or Sec.~\ref{subsec:implementation} for implementation
details and additional information). These include:

\begin{itemize}
\item \texttt{.active}: This attribute controls whether support for \texttt{smuggle}
  statements and onion comments is enabled (\texttt{True}) or
  disabled (\texttt{False}).  When Davos is first imported,
  the \texttt{.active} attribute is set to \texttt{True}.

\item \texttt{.auto\_rerun}: This attribute controls how
  Davos behaves when attempting to \texttt{smuggle} a new
  version of a package that was previously imported and cannot be
  reloaded. This can happen if the package includes extension modules
  that dynamically link C or C++ objects to the Python interpreter,
  and the code that generates those objects was changed between the
  previously imported and to-be-smuggled versions.  If this attribute
  is set to \texttt{True}, Davos will automatically restart
  the notebook kernel and rerun all code up to (and including) the
  current \texttt{smuggle} statement. If set to \texttt{False} (the default),
  Davos will instead issue a warning, pause execution, and
  prompt the user to either restart and rerun the notebook, or
  continue running with the previously imported package version until
  the next time the kernel is restarted manually.  Note that, as of
  this writing, the \texttt{.auto\_rerun} attribute is not supported
  in Google Colaboratory notebooks.

\item \texttt{.confirm\_install}: If set to \texttt{True} (default:
  \texttt{False}), Davos will require user confirmation
  before installing a smuggled package that does not yet exist in the
  user's environment.

\item \texttt{.pip\_executable}: This attribute's value specifies the
  path to the \texttt{pip} executable used to install smuggled
  packages. The default is programmatically determined from the Python
  environment and falls back to \texttt{sys.executable -m pip} if no
  executable can be found.

\item \texttt{.suppress\_stdout}: If this attribute is set to
  \texttt{True} (default: \texttt{False}), Davos suppresses
  printed (console) outputs from both itself and the installer program.
  This can be useful when smuggling packages that need to install many
  dependencies and/or generate extensive output. However, if the installer
  program throws an error, both its stdout and stderr streams will be
  displayed alongside the Python traceback to allow for debugging.

\item \texttt{.project}: \textcolor{red}{\textbf{TODO: fix this}} This attribute is a string that specifies the name of
the ``project'' associated with the current notebook. As described in
Section~\ref{subsec:projects}, a notebook's project determines where and how
any \texttt{smuggle}d dependencies are installed if they are not available in
the current runtime environment. By default, this attribute is named according
to the current notebook's absolute file path. However, the project name may be
customized to enable shared dependency installations across notebooks (see
Sec.~\ref{subsec:projects}).

@@

\subsubsection{Other top-level Davos functions}\label{subsec:toplevel}
\textcolor{red}{\textbf{TODO: add content}}

@@

IPython preprocesses all executed code as plain text before it is sent
to the Python compiler, in order to handle special constructs like
\texttt{\%magic} and \texttt{!shell} commands. Davos uses
this process to transform \texttt{smuggle} statements into
syntactically valid Python code. The Davos parser uses a
regular expression to match lines of code containing \texttt{smuggle}
statements (and, optionally, onion comments), extract relevant
information from their text, and replace them with equivalent calls to
the \texttt{smuggle()} function. For example, if a user runs a
notebook cell containing
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet6}
\end{center}
the code that is actually executed by the Python interpreter would be
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet7}
\end{center}
The call to the \texttt{smuggle()} function carries out
Davos's central logic by determining whether the smuggled
package must be installed, carrying out the installation if necessary,
and subsequently loading it into the namespace. This process is
outlined in Figure~\ref{fig:flow-chart}. Because the
\texttt{smuggle()} function is defined in the notebook namespace, it
is also possible (though never necessary) to call it
directly. Deactivating Davos will delete the name
``\texttt{smuggle}'' from the namespace, unless its value has been
overwritten and no longer refers to the \texttt{smuggle()}
function. It will also deregister the Davos parser from the
set of input transformers run when each notebook cell is
executed. While the overhead added by the Davos parser is
minimal, this may be useful, for example, when optimizing or precisely
profiling code.

>>>>

After importing Davos into a notebook, the top-level \texttt{davos} module provides a set of attributes whose values determine various aspects of Davos's behavior. A subset of these attributes are writeable options that may be modified to customize how, where, and when Davos installs smuggled packages (see Sec.~\ref{sec:illustrative-example} for an illustrative example and Sec.~\ref{subsec:implementation} for implementation details and additional information). These include:

%After importing Davos into a notebook, the top-level \texttt{davos} module provides a set of attributes whose values may be displayed interactively, checked programmatically at runtime, or modified to customize various aspects of Davos's behavior (see Sec.~\ref{sec:illustrative-example} for an illustrative example and Sec.~\ref{subsec:implementation} for implementation details and additional information). 
%These writeable attributes include:

%Davos's behavior may be customized by modifying a set of attributes attached to
%the \texttt{davos} module object that is added to the workspace when Davos is
%imported. These attributes may be modified, displayed, or checked
%programmatically at runtime (see Sec.~\ref{sec:illustrative-example} for an
%illustrative example or Sec.~\ref{subsec:implementation} for implementation
%details and additional information). These include:

\begin{itemize}
\item \texttt{.active}: This attribute controls whether support for \texttt{smuggle}
  statements and onion comments is enabled (\texttt{True}) or
  disabled (\texttt{False}).  When Davos is first imported,
  \texttt{davos.active} is set to \texttt{True}.

\item \texttt{.auto\_rerun}: This attribute controls how
  Davos behaves when attempting to \texttt{smuggle} a new
  version of a package that was previously loaded (via an \texttt{import} or \texttt{smuggle} statement) and cannot be
  reloaded. This can happen if the package includes extension modules
  that dynamically link C or C++ objects to the Python interpreter,
  and the code that generates those objects was changed between the
  previously loaded and to-be-smuggled versions.  If this attribute
  is set to \texttt{True}, Davos will automatically restart
  the notebook kernel and rerun all code up to (and including) the
  current \texttt{smuggle} statement. If set to \texttt{False} (the default),
  Davos will instead issue a warning, pause execution, and
  prompt the user to either restart and rerun the notebook, or
  continue running with the previously loaded package version until
  the next time the kernel is restarted manually.  Note that, as of
  this writing, setting \texttt{davos.auto\_rerun} to \texttt{True} is not 
  supported in Google Colaboratory notebooks.

\item \texttt{.confirm\_install}: If set to \texttt{True} (default:
  \texttt{False}), Davos will require user confirmation
  before installing a smuggled package that is not already
  available locally. This is primarily useful if the user has disabled 
  Davos's ``project'' system for isolating smuggled packages (see 
  Sec.~\ref{subsec:projects}) but still wants to carefully control what 
  packages are installed into their main Python environment.

\item \texttt{.pip\_executable}: This attribute's value specifies the
  path to the \texttt{pip} executable used to install smuggled
  packages. The default is programmatically determined from the user's Python
  environment and falls back to \texttt{<sys.executable> -m pip} if no
  executable can be found.

\item \texttt{.project}: This attribute's value is a \texttt{Project} instance representing the Davos project associated with the current notebook. As described in Section~\ref{subsec:projects}, Davos projects serve to isolate packages installed by \texttt{smuggle} statements from the user's main Python environment, and the \texttt{Project} class provides an interface for inspecting and managing projects at runtime. This attribute's default value is a notebook-specific project named for the absolute path to the notebook file. To change the project used in the current notebook (e.g., in order to share a project between multiple related notebooks), this attribute may be assigned a different \texttt{Project} instance or, for simplicity, the name of the desired project as a string or \texttt{pathlib.Path} (which will be converted to a \texttt{Project} on assignment). Alternatively, setting \texttt{davos.project} to \texttt{None} will disable project-based isolation for the current notebook and cause Davos to install packages directly into the main Python environment. This attribute is reset to its default value by the top-level \texttt{use\_default\_project()} function (see Sec.~\ref{subsec:toplevel}). For more information about Davos projects, see Section \ref{subsec:projects}.

\item \texttt{.suppress\_stdout}: If this attribute is set to
  \texttt{True} (default: \texttt{False}), Davos suppresses
  printed (console) outputs from both itself and the installer program.
  This can be useful when smuggling packages that require installing many
  dependencies and/or generate extensive output when built from source distributions. Note that if this option is enabled and the installer
  program throws an error, both its stdout and stderr streams will still be
  displayed alongside the Python traceback to allow for debugging.

@@

@@

\subsubsection{Other top-level Davos functions}\label{subsec:toplevel}
\textcolor{red}{\textbf{TODO: add content}}
%- .configure()
%- .get_project()
%- .prune_projects()
%- .require_pip()
%- .require_python()
%- .use_default_project()

@@

IPython preprocesses all executed code as plain text before it is sent
to the Python compiler, in order to handle special constructs like
\texttt{!}-prefixed shell commands and \texttt{\%}-prefixed ``magic'' commands. Davos uses
%``magic'' (i.e., \texttt{\%}-prefixed) and and `` \texttt{!shell} commands. Davos uses
this process to transform \texttt{smuggle} statements into
syntactically valid Python code. The Davos parser uses a
regular expression to match lines of code containing \texttt{smuggle}
statements (and, optionally, onion comments), extract relevant
information from their text, and replace them with equivalent calls to
the \texttt{smuggle()} function. For example, if a user runs a
notebook cell containing
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet6}
\end{center}
the code that is actually executed by the Python interpreter would be
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet7}
\end{center}
The call to the \texttt{smuggle()} function carries out
Davos's central logic by determining whether the smuggled
package must be installed, carrying out the installation if necessary,
and subsequently loading it into the namespace. This process is
outlined in Figure~\ref{fig:flow-chart}. Because the
\texttt{smuggle()} function is defined in the notebook namespace, it
is also possible (though never necessary) to call it
directly. Deactivating Davos will delete the name
``\texttt{smuggle}'' from the namespace, unless its value has been
overwritten and no longer refers to the \texttt{smuggle()}
function. It will also deregister the Davos parser from the
set of input transformers run when each notebook cell is
executed. While the overhead added by the Davos parser is
minimal, this may be useful, for example, when optimizing or precisely
profiling code.
